---
title: "MARTA Open Data"
output: html_notebook
---

```{r load R libraries}
library(ggplot2); library(dplyr); library(stringr); library(readr); library(lubridate)
```

### Get APC Data (North Avenue Corridor)
Source Link: http://opendata.itsmarta.com/hackathon/2017/February/APC/apcdata_NorthAve.csv

See [APC Data Dictionary](data_notes/apc-data-dictionary.md) for dataset info

```{r get_apc_data}
## Get apc data (North Avenue corridor) ----
apcdata <- read_csv("data_raw/opendata.itsmarta.com/apcdata_NorthAve.csv",
                    col_types = "ccccccccccccc") %>%
     mutate(calendar_day = mdy(calendar_day),
            route = as.integer(route),
            ons = as.integer(ons),
            offs = as.integer(offs),
            latitude = as.numeric(latitude),
            longitude = as.numeric(longitude))
```

### Get AVL Data
Source Links to files on opendata.itsmarta.com

* [avl_otpdata_week.csv](http://opendata.itsmarta.com/hackathon/2017/February/AVL_OTP/avl_otpdata_week.csv)
* [avl_otpdata_month.csv](http://opendata.itsmarta.com/hackathon/2017/February/AVL_OTP/avl_otpdata_month.csv)
* [avl_otpdata_year.csv](http://opendata.itsmarta.com/hackathon/2017/February/AVL_OTP/avl_otpdata_year.csv)

See [AVL Data Dictionary](data_notes/avl_otpdata-data-dictionary.md) for dataset info

```{r get_avl_data}
# ## Get avl data ----
# avl_week_raw <- read_csv("data_raw/opendata.itsmarta.com/avl_otpdata_week.csv", 
#                      col_types = paste(rep("c", times = 19), collapse = ""))
# # avl_month <- read_csv("data_raw/opendata.itsmarta.com/avl_otpdata_month.csv",
# #                       col_types = paste(rep("c", times = 19), collapse = ""))
# # avl_year <- read_csv("data_raw/opendata.itsmarta.com/avl_otpdata_year.csv",
# #                      col_types = paste(rep("c", times = 19), collapse = ""))
# 
# avl_week <- 
#      avl_week_raw %>%
#      mutate(calendar_day = dmy(calendar_day),
#             scheduled_time = as.integer(scheduled_time),
#             actual_arrival_time = as.integer(actual_arrival_time),
#             actual_depart_time = as.integer(actual_depart_time),
#             adherence_seconds = as.integer(adherence_seconds),
#             latitude = as.numeric(latitude) / 10000000,
#             longitude = as.numeric(longitude) / 10000000)
# 
# avl_week$adherence_category <- 
#      cut(avl_week$adherence_seconds, breaks = c(-10000, -300, 300, 10000))

```

### Parking Data
Source Links to files on opendata.itsmarta.com:

* [Parking_Info.csv](http://opendata.itsmarta.com/hackathon/2017/February/Parking_Data/Parking_Info.csv)
* [Station.csv](http://opendata.itsmarta.com/hackathon/2017/February/Parking_Data/Station.csv)

```{r get_parking_data}
parking <- 
     left_join(read_csv("data_raw/opendata.itsmarta.com/Parking_Info.csv", col_types = "iiiii_"),
               read_csv("data_raw/opendata.itsmarta.com/Station.csv", col_types = "ici"),
               by = "STATION_ID") %>%
     mutate(PERCENT_FILLED = ifelse(is.na(NUMBER_OF_CARS_PARKED) | is.na(NUMBER_OF_SPACES),
                                    NA, NUMBER_OF_CARS_PARKED / NUMBER_OF_SPACES))
parking
```

```{r Function: Prep AVL Data}
prep_avl_data <- function(timeframe){
if(timeframe == "week"){
     avldf_raw <- read_csv("data_raw/opendata.itsmarta.com/avl_otpdata_week.csv",
                           col_types = paste(rep("c", times = 19), collapse = ""))
} else if(timeframe == "month"){
     avldf_raw <- read_csv("data_raw/opendata.itsmarta.com/avl_otpdata_month.csv",
                           col_types = paste(rep("c", times = 19), collapse = ""))
} else if(timeframe == "year"){
     avldf_raw <- read_csv("data_raw/opendata.itsmarta.com/avl_otpdata_year.csv",
                           col_types = paste(rep("c", times = 19), collapse = ""))
}
avldf <- 
     avldf_raw %>%
     mutate(calendar_day = dmy(calendar_day),
            scheduled_time = as.integer(scheduled_time),
            actual_arrival_time = as.integer(actual_arrival_time),
            actual_depart_time = as.integer(actual_depart_time),
            adherence_seconds = as.integer(adherence_seconds),
            latitude = as.numeric(latitude) / 10000000,
            longitude = as.numeric(longitude) / 10000000)

avldf$adherence_category <- 
     cut(avldf$adherence_seconds, breaks = c(-10000, -300, 300, 10000))

longitude_group8 <- avldf %>% select(longitude) %>% distinct() %>% arrange(longitude) %>% mutate(longitude_group = ntile(longitude, n = 8))

latitude_group8 <- avldf %>% select(latitude) %>% distinct() %>% arrange(latitude) %>% mutate(latitude_group = ntile(latitude, n = 8))

avldf <- 
     avldf %>% 
     left_join(longitude_group8, by = "longitude") %>% 
     left_join(latitude_group8, by = "latitude")

return(avldf)
}

## Get AVL Data ----
# avl_week <- prep_avl_data("week")
# avl_month <- prep_avl_data("month")
avl_year <- prep_avl_data("year")
```

```{r Plot: adherence}
adherence_histogram <- function(df, plot_filename){

     p <- 
          ggplot() + 
          geom_bar(data = df, aes(adherence_seconds)) + 
          geom_vline(data = df, 
                     xintercept = quantile(df$adherence_seconds, 
                                           probs = c(.01, .99)), color = "red") + 
          geom_vline(data = df, 
                     xintercept = quantile(df$adherence_seconds, 
                                           probs = c(.1, .9)), color = "orange") +
          geom_vline(data = df, 
                     xintercept = quantile(df$adherence_seconds, 
                                           probs = c(.25, .75)), color = "green3") + 
          geom_vline(data = df, 
                     xintercept = quantile(df$adherence_seconds, 
                                           probs = c(.5)), color = "blue") + 
          geom_label(
               data = data_frame(
                    adherence_seconds = quantile(df$adherence_seconds, 
                                                 probs = c(.01, .1, .25, .5, .75, .9, .99)),
                    count = 575), 
               aes(adherence_seconds, count, label = round(adherence_seconds, 0))) + 
          ggtitle("AVL Data (week): Adherence Seconds", 
                  subtitle = "Colored lines show quantiles - Red: 1%/99%, Orange: 10%/90%, Green = 25%/75%, Blue = 50% (median)")
     
     # Side Effect: PNG file
     ggsave(filename = paste0("data_analysis/avl_adherence_histogram", plot_filename, ".png"),
            width = 10, height = 2.5)

     # Return: plot
     return(p)
}
```

```{r summarize avl data}
avl_year_summary_latlon <- 
     avl_year %>% group_by(longitude_group, latitude_group) %>% 
     summarise(adh_median = median(adherence_seconds), 
               adh_iqr = IQR(adherence_seconds),
               count = n(), 
               lon_min = min(longitude), lon_max = max(longitude), 
               lon_mid = (lon_min + lon_max) / 2,
               lat_min = min(latitude), lat_max = max(latitude),
               lat_mid = (lat_min + lat_max) / 2)

ggplot(avl_year_summary_latlon) + geom_point(aes(lon_mid, lat_mid, color = adh_median, size = count)) +
     scale_color_continuous(low = "red", high = "green") +
     scale_size_area()

ggplot(avl_year_summary_latlon) + geom_point(aes(count, adh_median, size = adh_iqr, color = )) + scale_size_area()
```

